---
title: "Dashboard Visualizations Example 1"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
---

# Introduction

Learn how to make Automated Dashboard with various correlation visualizations in R. First you need to install the `rmarkdown` package into your R library. Assuming that you installed the `rmarkdown`, next you create a new `rmarkdown` script in R. 

Reference

https://datascienceplus.com/automated-dashboard-with-various-correlation-visualizations-in-r/


```{r}
setwd("~/R/Complete") #change as needed

# options
options(echo=TRUE)
options(stringsAsFactors=FALSE)
options(scipen=999)  # turn-off scientific notation like 1e+48

if(!require(easypackages)){install.packages("easypackages")}
library(easypackages)
packages("flexdashboard", "tidyverse",  "ggalt", "plotly", "gganimate", "ggExtra", 
        "ggcorrplot", prompt = TRUE)

theme_set(theme_bw())  # pre-set the bw theme.
```


```{r}
data("midwest", package = "ggplot2")

midwest_select <- midwest %>% filter(poptotal > 350000, area < 0.1)

data(mpg, package="ggplot2")
g <- ggplot(mpg, aes(cty, hwy))
mpg_select <- mpg %>% filter(cty > 27)

data(mtcars)
```

# Dashboard

Row
-----------------------------------------------------------------------

### Chart A: Scatterplot

```{r warning=FALSE}
gg <- ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(aes(col=state, size=popdensity)) + 
  geom_smooth(method="loess", se=F) + 
  xlim(c(0, 0.1)) + 
  ylim(c(0, 500000)) + 
  labs(subtitle="Area Vs Population", 
       y="Population", 
       x="Area", 
       title="Scatterplot", 
       caption = "Source: midwest")

plot(gg)
ggplotly(p = ggplot2::last_plot())
```

### Chart B: Scatterplot + Encircle

```{r warning=FALSE}
ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(aes(col=state, size=popdensity)) +   # draw points
  geom_smooth(method="loess", se=F) + 
  xlim(c(0, 0.1)) + 
  ylim(c(0, 500000)) +   # draw smoothing line
  geom_encircle(aes(x=area, y=poptotal), 
                data=midwest_select, 
                color="red", 
                size=2, 
                expand=0.08) +   # encircle
  labs(subtitle="Area Vs Population", 
       y="Population", 
       x="Area", 
       title="Scatterplot + Encircle", 
       caption="Source: midwest")
```

Row
-----------------------------------------------------------------------

### Cart C: Jitter Plot

```{r}
g + geom_point() + 
  geom_smooth(method="lm", se=F) +
  labs(subtitle="mpg: city vs highway mileage", 
       y="hwy", 
       x="cty", 
       title="Scatterplot with overlapping points", 
       caption="Source: midwest")
ggplotly(p = ggplot2::last_plot())
```

### Cart D: Jitter Points

```{r}
# Scatterplot
theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(mpg, aes(cty, hwy))
g + geom_jitter(width = .5, size=1) +
  labs(subtitle="mpg: city vs highway mileage", 
       y="hwy", 
       x="cty", 
       title="Jittered Points")
ggplotly(p = ggplot2::last_plot())
```

Row
-----------------------------------------------------------------------

### Chart E: Counts Chart

```{r}
# Scatterplot
theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(mpg, aes(cty, hwy))
g + geom_count(col="tomato3", show.legend=F) +
  labs(subtitle="mpg: city vs highway mileage", 
       y="hwy", 
       x="cty", 
       title="Counts Plot")
ggplotly(p = ggplot2::last_plot())
```

### Chart F: Bubble plot

```{r}
# load package and data
data(mpg, package="ggplot2")
# mpg <- read.csv("http://goo.gl/uEeRGu")

mpg_select <- mpg[mpg$manufacturer %in% c("audi", "ford", "honda", "hyundai"), ]

# Scatterplot
theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(mpg_select, aes(displ, cty)) + 
  labs(subtitle="mpg: Displacement vs City Mileage",
       title="Bubble chart")

g + geom_jitter(aes(col=manufacturer, size=hwy)) + 
  geom_smooth(aes(col=manufacturer), method="lm", se=F)
ggplotly(p = ggplot2::last_plot())
```

Row
-----------------------------------------------------------------------
### Chart G: Marginal Histogram / Boxplot

```{r}
# Scatterplot
g <- ggplot(mpg, aes(cty, hwy)) + geom_count() + geom_smooth(method="lm", se=F)

ggMarginal(g, type = "histogram", fill="transparent")
ggMarginal(g, type = "boxplot", fill="transparent")
ggMarginal(g, type = "density", fill="transparent")
```

### Chart H: Correlogram
```{r}
# Correlation matrix
corr <- round(cor(mtcars), 1)

# Plot
ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of mtcars", 
           ggtheme=theme_bw)
```

