---
title: "Organizing R Shiny Apps"
output: html_document
---

# Introduction

Keeping R Shiny code organized can be a challenge. One method to organize your Shiny UI and Server code is to use a combination of R's list and source functions. 

- Difficulty Organizing Shiny UIs 
  - A UI Source Example That Usually Doesn't Work
  - When Source Fails in the UI
  - Solution
  - Usage Note
- Organizing the Shiny Server with Source
- Summary

# Difficulty Organizing Shiny UIs

If you have a Shiny application with a lot of controls, it's easy to get lost in the braces, parentheses and brackets. Using R's `list()` and `source()` functions, we can make UI code much more manageable by moving components out to other files.

To get us started, we are going to work with the starter app provided in R studio. Click File ???New ???Shiny Web App. . A dialogue box will pop up. Give the app a name, select *Single File App* and save to teh directory of your choice.

Start with the UI portion of the Shiny app. The UI can be broken down into two parts the *Side Bar Panel*  and the *Main Panel*" [lines 15-17*]. Factor out the content in the Side Bar Panel to a file named `Side_Panel_CTRL.`R and the Main Panel content to a file named `Main_Panel_OUTPUT.R`.

```{r eval=FALSE}
# Define UI for application that draws a histogram
ui <- fluidPage(

    # Application title
    titlePanel("Old Faithful Geyser Data"),

    # Sidebar with a slider input for number of bins 
    sidebarLayout(
        sidebarPanel(
            sliderInput("bins",
                        "Number of bins:", min = 1, max = 50, value = 30)
        ),

        # Show a plot of the generated distribution
        mainPanel(
           plotOutput("distPlot")
        )
    )
)
```

From the `app.R` file:

- MOVE the sliderInput to the `Side_Panel_Control.R` file
- MOVE the plotOutput to the `Main_Panel_OUTPUT.R`. 

Ensure the two new files are saved in the same folder as your `app.R` file.

Side_Panel_Control.R

```{r eval=FALSE}
sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30)
```

Main_Panel_OUTPUT.R

```{r eval=FALSE}
plotOutput("distPlot")
```

Once the files setup, add `source()` statements to  `app.R` file.

```{r eval=FALSE}
# Define UI for application that draws a histogram
ui <- fluidPage(
   titlePanel("List and Source UI"), # Title Page
 
   # Sidebar with a slider input for number of bins
   sidebarLayout(  
      sidebarPanel( # The Side Bar
        source(file = "Side_Panel_CTRL.R", local = T)[1]
      ),
   # Show a plot of the generated distribution    
   mainPanel( 
        source(file="Main_Panel_OUTPUT.R", local=T)[1]
      )
   )
)
```

> Without adding "[1]" to the end of the source() statement, the word TRUE pops up at the bottom of the slider control and the plot. Not sure why this happens but simply fixed.

# When Source Fails in the UI

If another control is added to `app.R` like a textbox control:

```{r eval=FALSE}
ui <- fluidPage(
    titlePanel("List and Source UI"), # Title Page
    
    # Sidebar with a slider input for number of bins
    sidebarLayout(  
        sidebarPanel( # The Side Bar
            source(file = "Side_Panel_CTRL.R", local = T)[1]
        ),
          #text box control
          textInput("textBox1", label = "Enter Value")   
    ),
        # Show a plot of the generated distribution    
        mainPanel( 
            source(file="Main_Panel_OUTPUT.R", local=T)[1]
        )
    )
)
```

The application works as intended.  BUT, if the textbox control is added to `Side_Panel_CTRL.R`, the app fails.  The need to organize the Shiny app requires a solution - and it is an easy one.

# Solution

The simple fix is to wrap the code in UI files with the list command.

SidePanel_CTRL.R

```{r eval=FALSE}
list(
  sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30)
  ),
  #text box control
          textInput("textBox1", label = "Enter Value")
)
```

Main_Panel_OUTOUT.R

```{r eval=FALSE}
list(
  plotOutput("distPlot")
)
```

`source()` and `list()` combo are particularly useful when working with tabbed panels.The controls for these two panels are coded in the source files.

```{r}
ui <- fluidPage(
    titlePanel("List and Source UI"), # Title Page
    
    # Sidebar with a slider input for number of bins
    sidebarLayout(  
        sidebarPanel( # The Side Bar
          tabsetPanel(
            tabPanel("Panel 1", source(file = "Side_Panel_CTRL.R", local = T)[1]),
            tabPanel("Panel 2", source(file = "Side_Panel2_CTRL.R", local = T)[1])
            )
        ),
        # Show a plot of the generated distribution    
        mainPanel( 
            source(file="Main_Panel_OUTPUT.R", local=T)[1]
        )
    )
)
```

# Organizing the Shiny Server

Server side Shiny is simply R-code and can be easily sourced from an external file using the `source()` command without list.

The only trick with server side Shiny is dealing with the reactive elements like `reactive({.})`. Keep `renderPlot({})` reactive element in the `app.R` and source from within:

**Original Server Code**
```{r eval=FALSE}
server <- function(input, output) {
   output$distPlot <- renderPlot({
     # generate bins based on input$bins from ui.R
     x    <- faithful[, 2] 
     bins <- seq(min(x), max(x), length.out = input$bins + 1)
     # draw the histogram with the specified number of bins
     hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
}
```

**New Server Code**
```{r eval=FALSE}
server <- function(input, output) {
   output$distPlot <- renderPlot({
     source(file="Server_Plot.R", local=T) 
   })
}
```

> Notice that compared to the UI side, we did not need to place the "[1]" at the end of the source command.

# Reference

http://r-bar.net/organize-r-shiny-list-source/